Variational Monte Carlo on Bosonic Systems

### 2018-04-06 project_1/article/article.tex
Spell check energy per particle section

    \subsection{Validity of Results}
        The non-interacting tests compared against the analytic "solution"
        showed to a very high degree the validity of the method. One can see
        this quite clearly from \autoref{fig:initial_problem_b}. Because we get
        results from simulations that correspond to the analytic results we can
        be fairly certain that the method can be expanded upon to include
        attributes outside the realm problems that are analytically solvable.
        The systems with elliptic traps and interacting particles are such
        systems.

        That said, the way to get the best confirmation of the results presented
        above is in comparison to experiment. This would be quite difficult
        as the systems presented are very theoretical in nature and most likely
        far from anything we would see in nature. Without being able to compare
        our result with anything happening in the real world this entire study would
        have no value except for the instrumental value, being no more than
        a theoretical exercise.  However, the one-body
        density is a measure of the role of many-body correlations which in
        some cases can be related to experiment. For a system of charged
        particles it can be used to determine a charge distribution, which in
        turn can be linked to experiment. It would be very fruitful to simulate a
        system of such particles that is more natural, i.e. real, but we realise the
        purpose of this study is first and foremost to introduce variational
        Monte Carlo as a tool in computational physics. We assume, however,
        that this study is easy enough to expand into a study of alkali atoms like
        $^{87}Rb$, $^{23}Na$ or $^7Li$.

    \subsection{Energy per particle}
        For the non-interacting harmonic oscillator system there is no
        difference between a new particle and new dimensions
        \footnote{This is not entirely true in the case of our implementation of
        the Metropolis algorithms as we draw a single random particle and move
        all dimensions instead of iterating over all particles and all
        dimensions.}. This means that the inclusion of a new particle or
        dimension yields a linear scaling in the energy. Thus if we compute the
        energy per particle for systems of varying size we expect to get the
        same energy. This is most easily seen in \autoref{fig:initial_problem_b}
        and in the expression for the exact energy \autoref{eq:exact_energy}.
        This, however, is \emph{not} the case in the interacting system due to
        the Jastrow factor. Looking at \autoref{tab:10_interacting} and
        \autoref{tab:100_interacting} we see that multiplying the expected
        energy of the former with $10$ yields a lower number than the latter. We
        can also see this behavior in \autoref{fig:interacting}, where the
        energy per particle for $N = \{10, 50, 100\}$ have been plotted in the
        same figure.

        \begin{figure}
            \includegraphics[width=244px]{../data/figures/problem_e.pdf}
            \caption{Local energies as a function of variational parameter
            $\alpha$ for different number of particles.}
            \label{fig:interacting}
        \end{figure}

        The reason for the increase in energy per particle can be ascribed to the
        repulsive interaction added by the Jastrow factor. The repulsive force can
        be explain thusly; The Jastrow factor  increases the value of the
        wavefunctions for any given particle if that particle is some distance away
         from any other particle. An increased value in the wavefucntion is equivalent
         to an increase in probability amplitude, again correpsonding to an increase in
         probability density, giving rise the the fact that particles are more likely to
         appear further from one another.

>>>>

    \subsection{Validity of Results}
        The non-interacting tests compared against the analytic "solution"
        showed to a very high degree the validity of the method. One can see
        this quite clearly from \autoref{fig:initial_problem_b}. Because we get
        results from simulations that correspond to the analytic results we can
        be fairly certain that the method can be expanded upon to include
        attributes outside the realm of problems that are analytically solvable.
        The systems with the elliptic traps and the interacting particles have
        just such attributes.

        That said, the way to get the best confirmation of the results presented
        above is in comparison to experiment. This would be quite difficult
        as the systems presented are very theoretical in nature and most likely
        far from anything we would see in nature. Without being able to compare
        our result with anything happening in the real world this entire study would
        have no value except for the instrumental value, being no more than
        a theoretical exercise.  However, the one-body
        density is a measure of the role of many-body correlations which in
        some cases can be related to experiment. For a system of charged
        particles it can be used to determine a charge distribution, which in
        turn can be linked to experiment. It would be very fruitful to simulate a
        system of such particles that is more natural, i.e. real, but we realise the
        purpose of this study is first and foremost to introduce variational
        Monte Carlo as a tool in computational physics. We assume, however, that
        this study is a good starting point for studies of alkali atoms like
        $^{87}$Rb, $^{23}$Na or $^7Li$.

    \subsection{Energy per particle}
        For the non-interacting harmonic oscillator system there is no
        difference between a new particle and a new dimension\footnote{This is
        not entirely true in the case of our implementation of the
        Metropolis-Hastings algorithm as we draw a single random particle and
        move all dimensions instead of iterating over all particles and all
        dimensions.}. This means that the inclusion of a new particle or
        dimension yields a linear scaling in the energy. Thus if we compute the
        energy per particle for systems of varying size we expect to get the
        same energy. This is most easily seen in \autoref{fig:initial_problem_b}
        and in the expression for the exact energy \autoref{eq:exact_energy}.

        This, however, is \emph{not} the case in the interacting system due to
        the Jastrow factor. Looking at \autoref{tab:10_interacting} and
        \autoref{tab:100_interacting} we see that multiplying the expected
        energy of the former with $10$ yields a lower number than the latter. We
        can also see this behavior in \autoref{fig:interacting}, where the
        energy per particle for $N = \{10, 50, 100\}$ have been plotted in the
        same figure.

        \begin{figure}
            \includegraphics[width=244px]{../data/figures/problem_e.pdf}
            \caption{Expected energy per particle as a function of variational
            parameter $\alpha$ for a varying number of particles.}
            \label{fig:interacting}
        \end{figure}

        The reason for the increase in energy per particle can be ascribed to
        the repulsive interaction added by the Jastrow factor. The repulsive
        force can be explained thusly: the Jastrow factor  increases the value
        of the local energy for any given particle if that particle is some
        distance away from any other particle. The closer the particles are, the
        more the local energy increases. The Jastrow factor will lower the
        probability for the particles to get too close to one another by
        reducing the probability density in the vicinity of other particles.
        However, if there are many particles in the system they will be closer
        to another particle and therefore increase the energy of the system.
